The Array of Long-Baseline Antennas for Taking Radio Observations from the Sub-Antarctic

### 2020-06-10 albatros_paper.tex
HCC working on text for single autonomous station

\subsection{Antenna}

@@

\subsection{Front-end active balun}\label{section:FEE}

@@

\subsection{Software and data acquisition}
\attention{Nivek?} \\
Text goes here.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{Figures/albatros_single_schematic/albatros_single_schematic.pdf}
    \caption{Single-antenna autonomous station block diagram. 
      A dual-polarization LWA
      antenna, equipped with an active balun (described above in \ref{section:FEE}), connects via 50m
      coaxial cables to the back-end readout electronics, 
      housed in a Faraday cage denoted by the dashed box. 
      The antenna signals are passed
      to a second-stage electronics chain consisting of filters and further amplfication.  The signals are
      digitized at 250~Msamp/s by a SNAP board, which includes an on-board FPGA that computes spectra from both antennas.  
      A Raspberry Pi controls the SNAP board, receives the spectrum of interest from the SNAP, and saves it as baseband data.}
    \label{Fig:Signal Chain}
  \end{center}
\end{figure}
\attention{Nivek?} \\

The full ALBATROS configuration will comprise an array of 10 autonomous antenna stations, recording the lowest 10-20MHz baseband to disk, to be collected periodically and subsequently correlated offline. The ALBATROS stations (located at the eight coastal hut sites plus the hydro shack and \prizm\ site) will
be separated by baselines of \SI{\approx {20}}{km} as shown in
\autoref{Fig:Marion}. One fully autonomous ALBATROS
station, shown in \autoref{Fig:autonomous}, was deployed in April 2019 at the hydro shack location (\ang{46;52.205;}S, \ang{37;50.612;}E) on Marion Island, as a first step in testing the technology needed to establish this array.

% HCC: we probably don't need these figures below, or we can
% incorporate them as subfigs into other places

% \begin{figure}[h]
% 	\begin{center}
% 		\includegraphics[width=0.5\linewidth]{Figures/balun.jpg}
% 		\caption{Unenclosed FEE mounted on the ALBATROS-EGG antenna supporting structure.} 
% 		\label{Fig:Balun}
% 	\end{center}
% \end{figure}
% 
% \begin{figure}[h]
% 	\begin{center}
% 		\includegraphics[width=0.7\linewidth]{Figures/Balun_Block.png}
% 		\caption{One Polarisation Block Diagram of the FEE \cite{2012PASP..124.1090H}}
% 		\label{Fig:Balun Schematic}
% 	\end{center}
% \end{figure}

\subsection{Readout electronics}
\attention{Nivek, Jon?} \\

The single antenna signal chain is shown in \autoref{Fig:Signal Chain}. 
% start of copied text
The back-end readout electronics are housed in a Faraday cage
located 50m away from the antennas. The analog signal chain consists 
a pair of high- and low-pass filters (Minicircuits
ZFHP-1R2+ and SLP-90+) that, as in the two-element configuration, together bandlimit the signal to
1.2--\SI{81}{MHz}. The filters are followed by a MiniCircuits ZFL-500+ amplifier.  The HPF and LPF contribute nominal insertion losses of 0.2dB and 0.14dB respectively, and the amplifier has a nominal gain of 25dB.

As with the dual-antenna configuration described above, a SNAP board digitizes the RF  at 250~Msamp/s. In this configuration, the SNAP board ADC is locked to a 10MHz reference produced by a Trimble Thunderbole E GPS disciplined clock module. The Raspberry Pi receives the spectrally decomposed baseband data \attention{which frequencies are we storing?} and stores it onto a set of hard drives. 
(\attention{reference the same frequency transform as in the 2-element case})
%end of copied text

\attention{Text here about one-bit stuff, pushing baseband through the RPi3,
clock requirements, data rates, etc.}

\subsection{Correlation}
While the basics of 1-bit correlation are well known in the limit where the signal level is much lower than the noise, ALBATROS may well be in the high-signal regime.  We present the basics of 1-bit correlation here, and show that even in the high-signal limit, 1-bit correlation remains viable.
The fundamental output of a 1-bit correlator for real data is $x_{ij} \equiv \left < \tilde{E}_i \tilde{E}_j \right >$.  $ \tilde{E}_i$ is the quantized version of the underlying electric field $E_i$ where $ \tilde{E}_i=1$ for $E_i>0$ and $ \tilde{E}_i=-1$ for $E_i<0$.  Note that complex data can be handled as the combination of real components.  This output is non-linear in the underlying true signal and noise levels, since the output saturates at unity for perfectly correlated electric fields.  To get at the true sky signals, we will need to undo this nonlinearity (the so-called Van Vleck corrections).  For a 2-level correlator, the expected output can be related to the true signals as follows (Van Vleck \& Middleton, e.g. D'Addario):
\begin{eqnarray}
\label{eqn:1bit_output}
\left < \sin(\frac{\pi}{2}x_{ij})\right > = \frac{V_{ij}}{\sqrt{V_{ij}+N_i}\sqrt{V_{ij}+N_j}}
\end{eqnarray}
where $V_{ij}$ is the true visibility, and $V_{ij}+N_{i,j}$ is the total noise power measured by antennas $i$ and $j$.  Inverting this relation to get the true signal requires knowing the noise powers, which is not possible from the 1-bit data themselves.  However, the SNAP board calculates these power levels on timescales of a few seconds, much faster than the power levels change, and so by combining the single-station power measurements with the cross-station cross correlation, we will be able to derive the true sky visibilities.

>>>>

\subsection{Antenna}\label{s:antenna}

@@

\subsection{Front-end active balun}\label{s:fee}

@@

\subsection{Software and data acquisition}

\attention{text goes here}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{Figures/albatros_single_schematic/albatros_single_schematic.pdf}
    \caption{Single-antenna autonomous station block diagram.  A
      dual-polarization LWA antenna, equipped with a front-end active
      balun, connects via 50-m coaxial cables to the back-end readout
      electronics, housed in a Faraday cage denoted by the dashed box.
      Each of the two antenna signals is passed to a second-stage
      electronics chain consisting of filters and further
      amplfication.  The signals are digitized at 250~Msamp/s by a
      SNAP board, which includes an on-board FPGA that computes
      channelized baseband and spectra from both inputs.  A Raspberry
      Pi controls the SNAP board and receives the baseband data and
      spectra.  The baseband data are saved to external hard drives.
      The system is powered by solar panels that charge a 24-V battery
      bank.}
    \label{Fig:albatros1_schem}
  \end{center}
\end{figure}

The full ALBATROS configuration will comprise an array of 10
autonomous antenna stations, each recording baseband over a tunable
10--20~MHz window within the full 0--125~MHz operating range.  The
baseband data will be collected periodically and subsequently
correlated offline.  The ALBATROS stations, located at the eight
coastal hut sites plus the hydro shack and \prizm\ site, will be
separated by baselines of $\sim20$~km as shown in
\autoref{Fig:marion_map_beam}. One fully autonomous ALBATROS station,
shown in \autoref{Fig:autonomous}, was deployed in April 2019 at the
hydro shack location (\ang{46;52.205;}S, \ang{37;50.612;}E) on Marion
Island as a first step in testing the technology needed to establish
the full array.  The system block diagram is shown in
\autoref{Fig:albatros1_schem}.  The antenna and front-end active balun
in the single autonomous station are identical to those used in the
two-element pathfinder (\S\ref{s:antenna} and \S\ref{s:fee}), and the
back-end electronics and power system are described in detail below.

% HCC: we probably don't need this figure
% \begin{figure}[h]
% 	\begin{center}
% 		\includegraphics[width=0.5\linewidth]{Figures/balun.jpg}
% 		\caption{Unenclosed FEE mounted on the ALBATROS-EGG antenna supporting structure.} 
% 		\label{Fig:Balun}
% 	\end{center}
% \end{figure}

The back-end readout electronics are housed in a Faraday cage located
50~m away from the antennas. The analog signal chain consists of a
pair of high- and low-pass filters (Minicircuits ZFHP-1R2+ and
SLP-150+) that together band-limit the signal to 1.2--140~MHz, and the
filters are followed by a MiniCircuits ZFL-500+ amplifier.  In
contrast to the two-element pathfinder, the low-pass cutoff is
increased from 81~MHz to 140~MHz to capture downlink signals at
137--138~MHz from the ORBCOMM satellite constellation.
\attention{words here about how we'll use orbcomm for backup timing
  information}

As with the two-element pathfinder, a SNAP board digitizes each of the
two RF signals at 250~Msamp/s. In the autonomous station
configuration, the SNAP board ADCs are locked to a 10-MHz reference
produced by a Trimble Thunderbolt~E GPS-disciplined clock module.  The
SNAP board FPGA computes two data products: 1) channelized baseband
data for each polarization over tunable frequency windows within the
0--125~MHz operating range, with the options of 1-, 2-, or 4-bit
compression, and 2) auto- and cross-spectra from the two polarizations
over the full 0--125~MHz span, accumulated over few-second intervals.
An RPi~3B+ controls the SNAP board and and receives the auto- and
cross-spectra via GPIO connections, and the spectra are saved to an
on-board SD card.  The baseband data are passed from the SNAP board to
the RPi via ethernet and written to external hard drives.  The
introduction of gigabit ethernet with the RPi~3B+ model has enabled
the high data throughput associated with writing baseband.  As a
benchmark, 1-bit baseband recording of two polarizations over 10~MHz
of bandwidth yields an approximate data rate of 5~MB/s, or 0.4~TB/day.

\subsection{Correlation}

While the basics of 1-bit correlation are well known in the limit where the signal level is much lower than the noise, ALBATROS may well be in the high-signal regime.  We present the basics of 1-bit correlation here, and show that even in the high-signal limit, 1-bit correlation remains viable.
The fundamental output of a 1-bit correlator for real data is $x_{ij} \equiv \left < \tilde{E}_i \tilde{E}_j \right >$.  $ \tilde{E}_i$ is the quantized version of the underlying electric field $E_i$ where $ \tilde{E}_i=1$ for $E_i>0$ and $ \tilde{E}_i=-1$ for $E_i<0$.  Note that complex data can be handled as the combination of real components.  This output is non-linear in the underlying true signal and noise levels, since the output saturates at unity for perfectly correlated electric fields.  To get at the true sky signals, we will need to undo this nonlinearity (the so-called Van Vleck corrections).  For a 2-level correlator, the expected output can be related to the true signals as follows (Van Vleck \& Middleton, e.g. D'Addario):
\begin{eqnarray}
\label{eqn:1bit_output}
\left < \sin(\frac{\pi}{2}x_{ij})\right > = \frac{V_{ij}}{\sqrt{V_{ij}+N_i}\sqrt{V_{ij}+N_j}}
\end{eqnarray}
where $V_{ij}$ is the true visibility, and $V_{ij}+N_{i,j}$ is the total noise power measured by antennas $i$ and $j$.  Inverting this relation to get the true signal requires knowing the noise powers, which is not possible from the 1-bit data themselves.  However, the SNAP board calculates these power levels on timescales of a few seconds, much faster than the power levels change, and so by combining the single-station power measurements with the cross-station cross correlation, we will be able to derive the true sky visibilities.
